---
title: "Prediction Model Writeup"
author: "Niyati Raghavan"
date: "09/13/2020"
output: html_document
---

### Overview

A model will be generated for the large amount of data about personal activity. The goal is to create a model with respect to the exercise performed using "classe" variable based on the training set.

### Data Processing

The training data for this project:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data for this project:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv 

The training and test datasets are loaded and further parsed such that the empty values and near zero variance are removed. 

```{r, echo=TRUE}
library("caret")
library(randomForest)

TrainingDataURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
TestDataURL  <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

training_data <- read.csv(url(TrainingDataURL), strip.white = TRUE, na.strings = c("NA",""))
test_data  <- read.csv(url(TestDataURL),  strip.white = TRUE, na.strings = c("NA",""))

training_data<-training_data[,colSums(is.na(training_data)) == 0]
test_data <-test_data[,colSums(is.na(test_data)) == 0]
```

### Model

Cross Validation is performed and the partition the training set into training and validation subsets. In general two partitions are created, with 70% and 30% for the original training set. The training parition will be used for modeling.

```{r, echo = TRUE}
training_data <- training_data[ , -c(1:6)]
test_data  <- test_data[ , -c(1:6)]

set.seed(10000)

# 70/30 model where, 70% to train and 30% to test out errors
partition_values <- createDataPartition(y=training_data$classe, p=0.70,list=FALSE)    
train_vals <- training_data[partition_values, ]
test_vals <- training_data[-partition_values, ] 

```

The Random Forest is used to model the training set/ Additionally, a confusion matrix is used to quantify the model.

Using this model, yields an overall accuracy is 99.85%. This is a highly accurate result indicates lower errors.

```{r, echo=TRUE}
model_rf <- randomForest(classe ~ ., data=train_vals, method="class", na.action=na.exclude)
predict_rf <- predict(model_rf, test_vals, type = "class")

confusionMatrix(test_vals$classe, predict_rf)
```

### Test Model

The model performed using random forest is used for the test data with the 20 entries.

```{r, echo=TRUE}
test_data_results <- predict(model_rf, test_data, type="class")
test_data_results
```
